results

### 2020-08-13 analysis/results.Rmd
variation

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
```

@@

Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio. Figure shows that the majority of journals supported the hybrid model. Overall, Elsevier published `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.

@@

flipped_jns_df <-
  els_jns %>%
    filter(issn %in% flipped_jns$issn)
```

@@

```{r}
els_jn_pub_volume <- els_pub_volume %>%
  filter(issn %in% eligible_jns$issn) %>%
  group_by(issued_year, issn) %>%
  summarise(articles = sum(articles),
            journals = n_distinct(issn))
hybrid_oa_per_jn <- hybrid_df %>%
  filter(oa_archive == FALSE) %>%
  group_by(issued_year, issn) %>%
  summarise(oa_articles = n())%>%
  inner_join(els_jn_pub_volume, by = c("issued_year", "issn")) %>%
  ungroup() %>%
  mutate(prop = oa_articles / articles * 100) %>% View()
```

The open access share varied across Elsevier hybrid journals. Since 2015, most hybrid journals have had a slow uptake rate of open access articles. In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.

>>>>

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(ggeconodist)
```

@@

Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio. Figure shows that the majority of journals were hybrid journals. Overall, Elsevier published `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.

@@

flipped_jns_df <- els_jns %>%
  filter(issn %in% flipped_jns$issn) %>%
  select(oa_model, issn, year)

hybrid_df <- left_join(hybrid_df, flipped_jns_df, by = c("issn", "issued_year" = "year")) %>%
  filter(!oa_model %in% "Open Access") %>%
  select(-oa_model)
```

@@

```{r}
els_jn_pub_volume <- els_pub_volume %>%
  filter(issn %in% eligible_jns$issn) %>%
  group_by(issued_year, issn) %>%
  summarise(articles = sum(articles),
            journals = n_distinct(issn))
hybrid_oa_per_jn <- hybrid_df %>%
  filter(oa_archive == FALSE) %>%
  group_by(issued_year, issn) %>%
  summarise(oa_articles = n())%>%
  inner_join(els_jn_pub_volume, by = c("issued_year", "issn")) %>%
  ungroup() %>%
  mutate(prop = oa_articles / articles) 
```

```{r}
perc_df <- hybrid_oa_per_jn %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_90 = quantile(prop, 0.9)) %>%
  ungroup()
```

The open access share varies across Elsevier hybrid journals. Since 2015, most hybrid journals have had a slow uptake rate of open access articles. Overall, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume. In 2019, the median open access share per hybrid journal was `r perc_df %>% filter(issued_year == 2019) %>% pull(median) * 100`%. In the same year, 90% of Elsevier hybrid journals had an open access uptake of `r perc_df %>% filter(issued_year == 2019) %>% pull(p_90) * 100`% or less.

```{r, fig.cap='Open access uptake in Elsevier journals per year in percent, visualised as diminutive distribution chart removing outliers. Since 2015, most hybrid journals have had a slow uptake rate of open access articles. In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.'}
a_plot <- perc_df %>%
ggplot() +
  geom_econodist(aes(x = factor(issued_year), ymin = p_10, median = median, ymax = p_90),
         stat = "identity", show.legend = TRUE) +
  scale_y_continuous(position = "right", 
                     labels = scales::percent_format(accuracy = 1L),
                    # expand = expansion(mult = c(0, 0.15)), 
                     limits = c(0,max(perc_df$p_90) + 0.01)) +
  coord_flip() +
  labs(
    x = NULL, y = NULL
   # title = "What is the uptake of hybrid open access among Elsevier journals?",
  #  subtitle = "Proportion of immediate open access articles published in subscription-based journals\nJan 2015 to July 2019",
   # caption = "Data source: Elsevier B.V. / Crossref"
  ) +
  theme_minimal(base_family = "Roboto") +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.border = element_blank())

grid.newpage()
left_align(a_plot, c("subtitle", "title", "caption")) %>% 
  add_econodist_legend(econodist_legend_grob(family = "Roboto"), below = "subtitle") %>% 
  grid.draw()  
```
